---
title: "ESPA Simulation and Real Data Application"
author: "Pratheepa Jeganathan, Alex Trindade"
date: "9/8/2018"
output: 
    html_document:
        toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(actuar)#     heavy tailed distributions
library(survival)
library(logspline) # Suffices to use logspline, the relevant ones in polspline are identical.
#library(polspline)
require(ggplot2)
require(gridExtra)
library(asaur)# dataset "gastricXelox"
```

##  loading necessary functions

ESPA PDF
```{r}
source("./Emp_SPA_pdf_t.R")
```

ESPA support
```{r}
source("./Emp_SPA_Dist_Supp.R")
```

ESPS survival
```{r}
source("./Emp_SPA_Survival_t.R")
```

##  Simulation

-   simulation from failure time ~ Weibull(shape=my.shape,scale=my.scale)
-   if model #4 ~ log-logistic(shape=my.shape,scale=my.scale) 
-   simulation from censoring time ~ exp(rate=1/my.scale)
-   % of censoring = target.cen
-   make the largest observation censored
-   Function Inputs:
    -   n,
    -   target censoring rate %,
    -   parameters of dist,
    -   model.num = model # (1-4) 
-   Returns: df.list, a list with 100 elements, each element of list is a rep, and is a data.frame with two columns (time, status).

```{r}
Simulate.Right.Censor.Data <- function(n, 
                                       target.cen, 
                                       my.shape, 
                                       my.scale, 
                                       model.num){
      set.seed(1000)
      m <- 1
      df.list <- list()
      while(m <= 100){# number of simulation reps
        t <- rweibull(n, shape=my.shape, scale=my.scale)
        if (model.num==4) t = rllogis(n, shape=my.shape)
        c <- rexp(n, rate=1/my.scale)
        #c <- rweibull(n, shape=my.shape, scale=my.scale) # to use the same censroing Y as T
        z <- pmin(t,c)
        time <- z
        status <- ifelse(t<=c,1,0)
        ## simulate <= target censoring rate and the largest obs is censored
        per.cen <- (n-sum(status))/n*100
        if(per.cen > target.cen|per.cen < (target.cen-5)|(status[which(time==max(time))]==1)){next}
        df.list[[m]] <- data.frame(time=time,status=status)
        m=m+1
      }
      return(df.list)
}
```

### Model 1: Weibull(shape=1,scale=10).

-   Hazard: h(t)=const.

```{r eval=FALSE}
out=Simulate.Right.Censor.Data(10,30,1,10,1)
saveRDS(out,"./simulated-data/df.list.model1_n10_c30.rds")
out=Simulate.Right.Censor.Data(10,60,1,10,1)
saveRDS(out,"./simulated-data/df.list.model1_n10_c60.rds")

out=Simulate.Right.Censor.Data(25,30,1,10,1)
saveRDS(out,"./simulated-data/df.list.model1_n25_c30.rds")
out=Simulate.Right.Censor.Data(25,60,1,10,1)
saveRDS(out,"./simulated-data/df.list.model1_n25_c60.rds")

out=Simulate.Right.Censor.Data(100,30,1,10,1)
saveRDS(out,"./simulated-data/df.list.model1_n100_c30.rds")
out=Simulate.Right.Censor.Data(100,60,1,10,1)
saveRDS(out,"./simulated-data/df.list.model1_n100_c60.rds")
```

### Model 2: Weibull(shape=1/3,scale=10)

-   Hazard: h(t) = monotone decreasing

```{r eval=FALSE}
out=Simulate.Right.Censor.Data(10,30,1/3,10,2)
saveRDS(out,"./simulated-data/df.list.model2_n10_c30.rds")
out=Simulate.Right.Censor.Data(10,60,1/3,10,2)
saveRDS(out,"./simulated-data/df.list.model2_n10_c60.rds")

out=Simulate.Right.Censor.Data(25,30,1/3,10,2)
saveRDS(out,"./simulated-data/df.list.model2_n25_c30.rds")
out=Simulate.Right.Censor.Data(25,60,1/3,10,2)
saveRDS(out,"./simulated-data/df.list.model2_n25_c60.rds")

out=Simulate.Right.Censor.Data(100,30,1/3,10,2)
saveRDS(out,"./simulated-data/df.list.model2_n100_c30.rds")
out=Simulate.Right.Censor.Data(100,60,1/3,10,2)
saveRDS(out,"./simulated-data/df.list.model2_n100_c60.rds")
```

### Model 3: Weibull(shape=2,scale=sqrt(2/3))

-   Hazard: h(t) = monotone increasing

```{r eval=FALSE}
out=Simulate.Right.Censor.Data(10,30,2,sqrt(2/3),3)
saveRDS(out,"./simulated-data/df.list.model3_n10_c30.rds")
out=Simulate.Right.Censor.Data(10,60,2,sqrt(2/3),3)
saveRDS(out,"./simulated-data/df.list.model3_n10_c60.rds")

out=Simulate.Right.Censor.Data(25,30,2,sqrt(2/3),3)
saveRDS(out,"./simulated-data/df.list.model3_n25_c30.rds")
out=Simulate.Right.Censor.Data(25,60,2,sqrt(2/3),3)
saveRDS(out,"./simulated-data/df.list.model3_n25_c60.rds")

out=Simulate.Right.Censor.Data(100,30,2,sqrt(2/3),3)
saveRDS(out,"./simulated-data/df.list.model3_n100_c30.rds")
out=Simulate.Right.Censor.Data(100,60,2,sqrt(2/3),3)
saveRDS(out,"./simulated-data/df.list.model3_n100_c60.rds")
```


### Model 4: Log-Logistic(shape=2, rate=1)

-   Hazard: h(t) = increasing then decreasing
```{r eval=FALSE}
out=Simulate.Right.Censor.Data(10,30,2,1,4)
saveRDS(out,"./simulated-data/df.list.model4_n10_c30.rds")
out=Simulate.Right.Censor.Data(10,60,2,1,4)
saveRDS(out,"./simulated-data/df.list.model4_n10_c60.rds")

out=Simulate.Right.Censor.Data(25,30,2,1,4)
saveRDS(out,"./simulated-data/df.list.model4_n25_c30.rds")
out=Simulate.Right.Censor.Data(25,60,2,1,4)
saveRDS(out,"./simulated-data/df.list.model4_n25_c60.rds")

out=Simulate.Right.Censor.Data(100,30,2,10,4)
saveRDS(out,"./simulated-data/df.list.model4_n100_c30.rds")
out=Simulate.Right.Censor.Data(100,60,2,10,4)
saveRDS(out,"./simulated-data/df.list.model4_n100_c60.rds")
```

##  Survival Function Plots

ESPA & logspline package approximate the survival functions of right-censored data.

We will also plots KM with 95% confidence bands.

-   Model 1: Weibull(shape=1,scale=10),     h(t)=const.
-   Model 2: Weibull(1/3,10),               h(t)=monotone dec.
-   Model 3: Weibull(2,sqrt{2/3}),          h(t)=monotone inc.
-   Model 4: Log-Logistic(shape=2,scale=1), h(t)=inc up to t=1 then dec

Data files are in "simulated-data" folder, m=100 reps from each model with (n=10,25) and censoring percentage 30%,60%. 

Each .rds data file is a list, each element of list corresponds to a rep, and is a data.frame with two columns (time, status). 

Check for example whether the largest observation in each rep is censored.
```{r}
df.list <- readRDS("./simulated-data/df.list.model1_n10_c30.rds")
xx=df.list[[3]]
which.order=sort(xx$time, index.return = TRUE)$ix
xx$time[which.order]
xx$status[which.order]
```

Define function to compute logsplines survival over a grid
-   INPUTS: 
    -   t.data: data.frame (col1=time, col2=status)
    -   t.grid: grid of quantiles where to compute surv
-   RETURNS:
    -   vector of survival probs over grid (if oldlogspline succeeded...)
    -   0 (if oldlogspline did NOT succeeded)
```{r}
logsplines.surv=function(t.data, 
                         t.grid){
      n=length(t.data$time)
      censored.obs = t.data$time[t.data$status==0]
      un.censored.obs = t.data$time[t.data$status==1]
      my.penalty=log(length(un.censored.obs))
      fit <- try(oldlogspline(uncensored=un.censored.obs, right=censored.obs,
                        lbound=0, penalty=my.penalty), silent=T)
      if (is.list(fit)){
          out=poldlogspline(t.grid, fit)
          return(1-out)
      }
      else return(0)
}
```


-   Function: plot true + logsplines + ESPA survival functions
    -   DEPENDS: on "logsplines.surv" function above.
-   INPUTS:
    -   t.data: data.frame (col1=time, col2=status)
    -   a=shape, b=scale (true Weibull params)
    -   model.num: Model # (for plot title)
    -   target.cen: censoring % (for plot title)
```{r}
plot.surv.fits <- function(a, 
                           b, 
                           t.data, 
                           model.num, 
                           target.cen, 
                           print.legend){
      eps=10^(-4); n=length(t.data$time); blue.is.ok=TRUE
      time=t.data$time; status=t.data$status
      # KM fit
      survob=Surv(time=time,event=status,type='right')
      fit=survfit(survob~1,conf.type="plain")
      
       # ESPA & true survivals
      source('./Emp_SPA_Dist_Supp.R')
      supp=Emp_SPA_Dist_Supp(time,status)
      my.grid=seq(supp[1],supp[2],by=(supp[2]-supp[1])/100)
     
      source('./Emp_SPA_Survival_t.R')
      spa.surv=rep(0,length(my.grid)); tru.surv=rep(0,length(my.grid))
      
      for(i in 1:length(my.grid)){
        spa.surv[i] = Emp_SPA_Survival_t(my.grid[i],time,status)
        tru.surv[i] = 1-pweibull(my.grid[i], shape=a, scale=b)
        if (model.num==4) tru.surv[i] = pllogis(my.grid[i], shape=2, lower.tail=F)
      }
      
      # get logsplines surv
      spl.surv=logsplines.surv(t.data, my.grid)
      # flag when logsplines did NOT succeed
      if (length(spl.surv)==1) blue.is.ok=FALSE
      # refine grid to be above epsilon
      new.idx=seq(1,length(my.grid))[tru.surv>eps];
      my.grid=my.grid[new.idx]; t.upper=max(my.grid)
      spa.surv=spa.surv[new.idx]
      tru.surv=tru.surv[new.idx]
      if (blue.is.ok) spl.surv=spl.surv[new.idx]
      # plot KM and overlay remianing ones
      plot(fit,lty=c(1,3,3),main='',xlab="time",conf.int=T, ylab="",
         yaxs="i",lwd=2,xlim=c(0,t.upper), col=c("grey","grey","grey"))
      points(my.grid,spa.surv,lwd=2,type='l',col='red',lty=2)
      points(my.grid,tru.surv,lwd=2,type='l',col='black',lty=1)
      if (blue.is.ok) points(my.grid,spl.surv,lwd=2,type='l',col='blue',lty=2)
      if (print.legend) legend('topright',legend=c('KM','ESPA','true','Logsplines'),
                               col=c("grey",'red','black','blue'), lty=c(1,2,1,2))
      title(main=bquote(paste('Survival: Model ',.(model.num),', n=',.(n),
                              ', (',.(target.cen),'% censor)')))
}
```

Analyze runs from Model 1
```{r fig.width=15,fig.height=15}
my.shape=1; my.scale=10; my.model=1; my.censor=30
df.list <- readRDS("./simulated-data/df.list.model1_n10_c30.rds")
par(mfrow=c(3,3))
for (i in 1:9){
  plot.surv.fits(my.shape, my.scale, df.list[[i]], my.model, my.censor, TRUE)
}
par(mfrow=c(1,1))
#
dev.off()
```

Analyze runs from Model 2
```{r fig.width=15,fig.height=15}
my.shape=1/3; my.scale=10; my.model=2; my.censor=60
df.list <- readRDS("./simulated-data/df.list.model2_n25_c60.rds")
par(mfrow=c(3,3))
for (i in c(1:6,8:10)){
  plot.surv.fits(my.shape, my.scale, df.list[[i]], my.model, my.censor, TRUE)
}
par(mfrow=c(1,1))
#
dev.off()
```

Analyze runs from Model 3
```{r fig.width=15,fig.height=15}
my.shape=2; my.scale=sqrt(2/3); my.model=3; my.censor=30
df.list <- readRDS("./simulated-data/df.list.model3_n25_c30.rds")
par(mfrow=c(3,3))
for (i in 1:9){
  plot.surv.fits(my.shape, my.scale, df.list[[i]], my.model, my.censor, TRUE)
}
par(mfrow=c(1,1))
#
dev.off()
```

Analyze runs from Model 4
```{r fig.width=15,fig.height=15}
my.shape=2; my.scale=1; my.model=4; my.censor=60
df.list <- readRDS("./simulated-data/df.list.model4_n25_c60.rds")
par(mfrow=c(3,3))
for (i in 1:9){
  plot.surv.fits(my.shape, my.scale, df.list[[i]], my.model, my.censor, TRUE)
}
par(mfrow=c(1,1))
#
dev.off()
```

###  Survival Plots

####    Analyze runs from Models 1-4 (all in one), n=10, 30% censored
```{r fig.width=15,fig.height=15}
pdf(file="Plots/Model-all-n10-c30.pdf", pointsize=12, paper="a4r",width=0,height=0)
# Mod 1
my.shape=1; my.scale=10; my.model=1; my.censor=30;
df.list <- readRDS("./simulated-data/df.list.model1_n10_c30.rds")
par(mfrow=c(4,4))
for (i in 1:4){
  plot.surv.fits(my.shape, my.scale, df.list[[i]], my.model, my.censor, FALSE)
}
# Mod 2
my.shape=1/3; my.scale=10; my.model=2;
df.list <- readRDS("./simulated-data/df.list.model2_n10_c30.rds")
for (i in 1:4){
  plot.surv.fits(my.shape, my.scale, df.list[[i]], my.model, my.censor, FALSE)
}
# Mod 3
my.shape=2; my.scale=sqrt(2/3); my.model=3;
df.list <- readRDS("./simulated-data/df.list.model3_n10_c30.rds")
for (i in 1:4){
  plot.surv.fits(my.shape, my.scale, df.list[[i]], my.model, my.censor, FALSE)
}
# Mod 4
my.shape=2; my.scale=1; my.model=4;
df.list <- readRDS("./simulated-data/df.list.model4_n10_c30.rds")
for (i in 1:4){
  plot.surv.fits(my.shape, my.scale, df.list[[i]], my.model, my.censor, FALSE)
}
par(mfrow=c(1,1))
#
dev.off()

```

####    Analyze runs from Models 1-4 (all in one), n=25, 30% censored
```{r fig.width=15,fig.height=15}
pdf(file="Plots/Model-all-n25-c30.pdf", pointsize=12, paper="a4r",width=0,height=0)
# Mod 1
my.shape=1; my.scale=10; my.model=1; my.censor=30;
df.list <- readRDS("./simulated-data/df.list.model1_n25_c30.rds")
par(mfrow=c(4,4))
plot.surv.fits(my.shape, my.scale, df.list[[1]], my.model, my.censor, TRUE)
for (i in 2:4){
  plot.surv.fits(my.shape, my.scale, df.list[[i]], my.model, my.censor, FALSE)
}
# Mod 2
my.shape=1/3; my.scale=10; my.model=2;
df.list <- readRDS("./simulated-data/df.list.model2_n25_c30.rds")
for (i in 1:4){
  plot.surv.fits(my.shape, my.scale, df.list[[i]], my.model, my.censor, FALSE)
}
# Mod 3
my.shape=2; my.scale=sqrt(2/3); my.model=3;
df.list <- readRDS("./simulated-data/df.list.model3_n25_c30.rds")
for (i in 1:4){
  plot.surv.fits(my.shape, my.scale, df.list[[i]], my.model, my.censor, FALSE)
}
# Mod 4
my.shape=2; my.scale=1; my.model=4;
df.list <- readRDS("./simulated-data/df.list.model4_n25_c30.rds")
for (i in 1:4){
  plot.surv.fits(my.shape, my.scale, df.list[[i]], my.model, my.censor, FALSE)
}
par(mfrow=c(1,1))
#
dev.off()
```

####    Analyze runs from Models 1-4 (all in one), n=10, 60% censored

```{r fig.width=15,fig.height=15}
pdf(file="Plots/Model-all-n10-c60.pdf", pointsize=12, paper="a4r",width=0,height=0)
# Mod 1
my.shape=1; my.scale=10; my.model=1; my.censor=60;
df.list <- readRDS("./simulated-data/df.list.model1_n10_c60.rds")
par(mfrow=c(4,4))
plot.surv.fits(my.shape, my.scale, df.list[[1]], my.model, my.censor, TRUE)
for (i in 2:4){
  plot.surv.fits(my.shape, my.scale, df.list[[i]], my.model, my.censor, FALSE)
}
# Mod 2
my.shape=1/3; my.scale=10; my.model=2;
df.list <- readRDS("./simulated-data/df.list.model2_n10_c60.rds")
for (i in 1:4){
  plot.surv.fits(my.shape, my.scale, df.list[[i]], my.model, my.censor, FALSE)
}
# Mod 3
my.shape=2; my.scale=sqrt(2/3); my.model=3;
df.list <- readRDS("./simulated-data/df.list.model3_n10_c60.rds")
for (i in 1:4){
  plot.surv.fits(my.shape, my.scale, df.list[[i]], my.model, my.censor, FALSE)
}
# Mod 4
my.shape=2; my.scale=1; my.model=4;
df.list <- readRDS("./simulated-data/df.list.model4_n10_c60.rds")
for (i in 1:4){
  plot.surv.fits(my.shape, my.scale, df.list[[i]], my.model, my.censor, FALSE)
}
par(mfrow=c(1,1))
#
dev.off()

```

####    Analyze runs from Models 1-4 (all in one), n=25, 60% censored
```{r fig.width=15,fig.height=15}
pdf(file="Plots/Model-all-n25-c60.pdf", pointsize=12, paper="a4r",width=0,height=0)
# Mod 1
my.shape=1; my.scale=10; my.model=1; my.censor=60;
df.list <- readRDS("./simulated-data/df.list.model1_n25_c60.rds")
par(mfrow=c(4,4))
plot.surv.fits(my.shape, my.scale, df.list[[1]], my.model, my.censor, TRUE)
for (i in 2:4){
  plot.surv.fits(my.shape, my.scale, df.list[[i]], my.model, my.censor, FALSE)
}
# Mod 2
my.shape=1/3; my.scale=10; my.model=2; my.censor=60;
df.list <- readRDS("./simulated-data/df.list.model2_n25_c60.rds")
for (i in 1:4){
  plot.surv.fits(my.shape, my.scale, df.list[[i]], my.model, my.censor, FALSE)
}
# Mod 3
my.shape=2; my.scale=sqrt(2/3); my.model=3; my.censor=60;
df.list <- readRDS("./simulated-data/df.list.model3_n25_c60.rds")
for (i in 1:4){
  plot.surv.fits(my.shape, my.scale, df.list[[i]], my.model, my.censor, FALSE)
}
# Mod 4
my.shape=2; my.scale=1; my.model=4;
df.list <- readRDS("./simulated-data/df.list.model4_n25_c60.rds")
for (i in 1:4){
  plot.surv.fits(my.shape, my.scale, df.list[[i]], my.model, my.censor, FALSE)
}
par(mfrow=c(1,1))
#
dev.off()
```

##  Integrated Squared Errors of survival approximations 

Figures 1, 3, 4, and 5 in the paper

Function to produce ggplot2 plots of ISEs.
    -   It branches depending on whether there are NA's in the spline ISEs, in which case it only plots the difference (spline-espa). 
    -   INPUTS:
        -   b.dat    = dataframe returned by Compute.Model.ISEs 
        -   my.label = label to print on plot LHS
    -   OUTPUT: out.box & out.den (Boxplot & PDF plot objects)

```{r}
Get.Box.Pdf.Plots <- function(b.dat, my.label){
  case1=FALSE; case2=FALSE; case3=FALSE; 
  idx=which(is.na(b.dat$spline)); m=dim(b.dat)[1]; my.title.size=13
  if (length(idx)==0) case3=TRUE # no NA's -> plot espa & splines separately
  if (length(idx)==m) case1=TRUE # all NA's -> plot only espa
  if (length(idx)>0 && length(idx)<m) case2=TRUE # some NA's -> plot difference
  if (case1){
#    my.label=paste(my.label,": \n 100% of spline method had no solution")
    ise=log(b.dat$espa)
    Method=factor(rep("espa",m))
    df2=data.frame(Method, ise)
    bg2 <- ggplot(df2, aes(Method, ise))
    bg2 <- bg2 + geom_boxplot(aes(fill = Method)) +
       xlab("") + ylab("") + ggtitle(my.label)
    bg2 <- bg2 + scale_fill_manual(values="#E69F00")
    dg2 <- ggplot(df2, aes(x=ise,colour=Method,group=Method))
    dg2 <- dg2 + xlab("log(ISE)") + geom_density(fill="#E69F00")+ylab("")
    dg2 <- dg2 + scale_color_manual(values="#E69F00") + ggtitle(my.label)+ 
           theme(plot.title = element_text(size=my.title.size, face = "bold"))
  }
  if (case2){
    b.dat=b.dat[-idx,]; m=dim(b.dat)[1]; my.per=100-m
#    my.label=paste(my.label,": \n ",my.per,"% of spline method had no solution")
    ise=log(b.dat$spline)-log(b.dat$espa)
    Method=factor(rep("Difference \n (spline-espa)",m))
    df2=data.frame(Method, ise)
    bg2 <- ggplot(df2, aes(Method, ise))
    bg2 <- bg2 + geom_boxplot(aes(fill = Method)) +
       xlab("") + ylab("") + ggtitle(my.label)
    bg2 <- bg2 + scale_fill_manual(values="#56B4E9")
    dg2 <- ggplot(df2, aes(x=ise, colour=Method, group=Method))
    dg2 <- dg2 + xlab("log(ISE)") + geom_density(fill="#56B4E9")+ylab("")
    dg2 <- dg2 + scale_color_manual(values="#56B4E9")+ ggtitle(my.label)+
           theme(plot.title = element_text(size=my.title.size, face = "bold"))
  }
  if (case3){
    ise=log(c(b.dat$espa,b.dat$spline))
    Method=factor(c(rep("espa",m),rep("spline",m)))
    df2=data.frame(Method, ise)
    bg2 <- ggplot(df2, aes(Method, ise))
    bg2 <- bg2 + geom_boxplot(aes(fill = Method)) +
         xlab("") + ylab("") + ggtitle(my.label) 
    dg2 <- ggplot(df2, aes(x=ise, colour=Method, group=Method, linetype=Method))
    dg2 <- dg2 + xlab("log(ISE)") + geom_density(fill=NA)+ ggtitle(my.label)+
           theme(plot.title = element_text(size=my.title.size, face = "bold"))+ylab("")
   }
   out=list("box"=bg2, "den"=dg2)
   return(out)
}
```

-   Function: Computes ISEs for logsplines & ESPA survival ftns
-   INPUTS:
    -   df.list: list of simulated data, each element is data.frame (col1=time, col2=status)
    -   my.shape=shape, my.scale=scale (true Weibull params)
    -   model.num: model # (1-4)
-   RETURNS: data.frame with ISEs in each col (espa, spline, KM)
-   The function has sub-functions
    (1) true survival function
    (2) squared error of ESPA survival 
    (3) squared error of spline survival approximation
    (4) squared error KM 
    

```{r }
Compute.Model.ISEs <- function(df.list, 
                               my.shape, 
                               my.scale, 
                               model.num){
    source('./Emp_SPA_Dist_Supp.R')
    source('./Emp_SPA_Survival_t.R')

#   true survival function
    t.fun=function(t,shape,scale) 1-pweibull(t,shape=shape,scale=scale)
    if (model.num==4){
        t.fun=function(t,shape,scale) pllogis(t, shape=shape, lower.tail=F)
    }
    
#   squared error of ESPA survival function
    sq.error.spa=function(x,time,status){
        len.x=length(x); sq.error=numeric(0)
        for(i in 1:len.x){
                true.den=t.fun(x[i],my.shape,my.scale)
                saddle.den=Emp_SPA_Survival_t(x[i],time,status)
                sq.error[i]=(true.den-saddle.den)^2       
        }
        return(sq.error)
    }

#   squared error of spline survival function
    sq.error.spl=function(x,time,status){
        true.den=t.fun(x,my.shape,my.scale); len.x=length(x)
        censored.obs = time[status==0]; un.censored.obs = time[status==1]
        my.penalty=log(length(un.censored.obs))
        fit <- try(oldlogspline(uncensored=un.censored.obs, right=censored.obs,
                        lbound=0, penalty=my.penalty), silent=T)
        if (is.list(fit)){
          spline.den=1-poldlogspline(x, fit)
          out=(true.den-spline.den)^2
        } else out=rep(NA,len.x)
        return(out)
    }
    
#   squared error KM 
    sq.error.km=function(x,time,status){
        true.den=t.fun(x,my.shape,my.scale); len.x=length(x)
        survob=Surv(time=time,event=status,type='right')
        fit=survfit(survob~1,conf.type="plain")
        fit.km <- stepfun(fit$time, c(1, fit$surv))
        km.den=fit.km(x)
        out=(true.den-km.den)^2
        return(out)
    }

#   Compute ISEs

    m=length(df.list); ise.spa=rep(0,m); ise.spl=rep(0,m); ise.km=rep(0,m); 
        for (i in 1:m){
            t.data = df.list[[i]]
            time = t.data$time; status=t.data$status
            supp = Emp_SPA_Dist_Supp(time,status); 
            if(supp[1]==0){
                supp[1] = supp[1]+.01
                }
            
            ise.spa[i] = tryCatch(integrate(sq.error.spa, supp[1], supp[2], time=time, status=status, abs.tol=.1)$value, error=function(e) NA)
            
            if (is.na(ise.spa[i])){
              ise.spa[i]=tryCatch(integrate(sq.error.spa, supp[1], supp[2], time=time, status=status, abs.tol=.2)$value, error=function(e) NA)
            }
            
            ise.spl[i]=tryCatch(integrate(sq.error.spl, supp[1], supp[2], time=time, status=status, abs.tol=.1)$value, error=function(e) NA) 
            
            ise.km[i]=tryCatch(integrate(sq.error.km, supp[1], supp[2], time=time, status=status, abs.tol=.1)$value, error=function(e) NA) 
        }
  
    out=data.frame(espa=ise.spa,spline=ise.spl,km=ise.km)
  
    return(out)
}
```

### Model 1 survival function log(ISE) values for the ESPA and SPLINE estimators 

```{r eval=FALSE}
a=1; b=10
a.list <- readRDS("./simulated-data/df.list.model1_n10_c30.rds")
ise.mod1.n10.c30=Compute.Model.ISEs(a.list, a, b, 1)
a.list <- readRDS("./simulated-data/df.list.model1_n10_c60.rds")
ise.mod1.n10.c60=Compute.Model.ISEs(a.list, a, b, 1)
a.list <- readRDS("./simulated-data/df.list.model1_n25_c30.rds")
ise.mod1.n25.c30=Compute.Model.ISEs(a.list, a, b, 1)
a.list <- readRDS("./simulated-data/df.list.model1_n25_c60.rds")
ise.mod1.n25.c60=Compute.Model.ISEs(a.list, a, b, 1)
a.list <- readRDS("./simulated-data/df.list.model1_n100_c30.rds")
ise.mod1.n100.c30=Compute.Model.ISEs(a.list, a, b, 1)
a.list <- readRDS("./simulated-data/df.list.model1_n100_c60.rds")
ise.mod1.n100.c60=Compute.Model.ISEs(a.list, a, b, 1)
save.image(file="./ISEs/Model1-ISEs.RData")
```

```{r fig.width=15,fig.height=15}
load(file="./ISEs/Model1-ISEs.RData")

out1=Get.Box.Pdf.Plots(ise.mod1.n10.c30,"n=10, c=30% (m=100)")
out2=Get.Box.Pdf.Plots(ise.mod1.n10.c60,"n=10, c=60% (m=100)")
out3=Get.Box.Pdf.Plots(ise.mod1.n25.c30,"n=25, c=30% (m=100)")
out4=Get.Box.Pdf.Plots(ise.mod1.n25.c60,"n=25, c=60% (m=100)")
out5=Get.Box.Pdf.Plots(ise.mod1.n100.c30,"n=100, c=30% (m=100)")
out6=Get.Box.Pdf.Plots(ise.mod1.n100.c60,"n=100, c=60% (m=97)")
## plot all
pdf(file="./Plots/Model1-ISEs-all.pdf", pointsize=12, paper="a4r",width=0,height=0)
grid.arrange(out1$den, out2$den, out3$den, out4$den, out5$den, out6$den, ncol=3, nrow=2)
dev.off()
```

### Model 2 survival function log(ISE) values for the ESPA and SPLINE estimators 
```{r eval=FALSE}
a=1/3; b=10
a.list <- readRDS("./simulated-data/df.list.model2_n10_c30.rds")
ise.mod2.n10.c30=Compute.Model.ISEs(a.list, a, b, 2)
a.list <- readRDS("./simulated-data/df.list.model2_n10_c60.rds")
ise.mod2.n10.c60=Compute.Model.ISEs(a.list, a, b, 2)
a.list <- readRDS("./simulated-data/df.list.model2_n25_c30.rds")
ise.mod2.n25.c30=Compute.Model.ISEs(a.list, a, b, 2)
a.list <- readRDS("./simulated-data/df.list.model2_n25_c60.rds")
ise.mod2.n25.c60=Compute.Model.ISEs(a.list, a, b, 2)
a.list <- readRDS("./simulated-data/df.list.model2_n100_c30.rds")
ise.mod2.n100.c30=Compute.Model.ISEs(a.list, a, b, 2)
a.list <- readRDS("./simulated-data/df.list.model2_n100_c60.rds")
ise.mod2.n100.c60=Compute.Model.ISEs(a.list, a, b, 2)
save.image(file="./ISEs/Model2-ISEs.RData")
```

```{r fig.width=15,fig.height=15}
load(file="./ISEs/Model2-ISEs.RData")

out1=Get.Box.Pdf.Plots(ise.mod2.n10.c30,"n=10, c=30% (m=100)")
out2=Get.Box.Pdf.Plots(ise.mod2.n10.c60,"n=10, c=60% (m=100)")
out3=Get.Box.Pdf.Plots(ise.mod2.n25.c30,"n=25, c=30% (m=100)")
out4=Get.Box.Pdf.Plots(ise.mod2.n25.c60,"n=25, c=60% (m=35)")
out5=Get.Box.Pdf.Plots(ise.mod2.n100.c30,"n=100, c=30% (m=83)")
out6=Get.Box.Pdf.Plots(ise.mod2.n100.c60,"n=100, c=60% (m=41)")
## plot all
pdf(file="./Plots/Model2-ISEs-all.pdf", pointsize=12, paper="a4r",width=0,height=0)
grid.arrange(out1$den, out2$den, out3$den, out4$den, out5$den, out6$den, ncol=3, nrow=2)
dev.off()
```

### Model 3 survival function log(ISE) values for the ESPA and SPLINE estimators 
```{r eval=FALSE}
a=2; b=sqrt(2/3)
a.list <- readRDS("./simulated-data/df.list.model3_n10_c30.rds")
ise.mod3.n10.c30=Compute.Model.ISEs(a.list, a, b, 3)
a.list <- readRDS("./simulated-data/df.list.model3_n10_c60.rds")
ise.mod3.n10.c60=Compute.Model.ISEs(a.list, a, b, 3)
a.list <- readRDS("./simulated-data/df.list.model3_n25_c30.rds")
ise.mod3.n25.c30=Compute.Model.ISEs(a.list, a, b, 3)
a.list <- readRDS("./simulated-data/df.list.model3_n25_c60.rds")
ise.mod3.n25.c60=Compute.Model.ISEs(a.list, a, b, 3)
a.list <- readRDS("./simulated-data/df.list.model3_n100_c30.rds")
ise.mod3.n100.c30=Compute.Model.ISEs(a.list, a, b, 3)
a.list <- readRDS("./simulated-data/df.list.model3_n100_c60.rds")
ise.mod3.n100.c60=Compute.Model.ISEs(a.list, a, b, 3)
save.image(file="./ISEs/Model3-ISEs.RData")
```

```{r fig.width=15,fig.height=15}
load(file="./ISEs/Model3-ISEs.RData")

out1=Get.Box.Pdf.Plots(ise.mod3.n10.c30,"n=10, c=30% (m=100)")
out2=Get.Box.Pdf.Plots(ise.mod3.n10.c60,"n=10, c=60% (m=100)")
out3=Get.Box.Pdf.Plots(ise.mod3.n25.c30,"n=25, c=30% (m=100)")
out4=Get.Box.Pdf.Plots(ise.mod3.n25.c60,"n=25, c=60% (m=100)")
out5=Get.Box.Pdf.Plots(ise.mod3.n100.c30,"n=100, c=30% (m=100)")
out6=Get.Box.Pdf.Plots(ise.mod3.n100.c60,"n=100, c=60% (m=100)")
## plot all
pdf(file="./Plots/Model3-ISEs-all.pdf", pointsize=12, paper="a4r",width=0,height=0)
grid.arrange(out1$den, out2$den, out3$den, out4$den, out5$den, out6$den, ncol=3, nrow=2)
dev.off()
```

### Model 4 survival function log(ISE) values for the ESPA and SPLINE estimators
```{r eval=FALSE}
a=2; b=1
a.list <- readRDS("./simulated-data/df.list.model4_n10_c30.rds")
ise.mod4.n10.c30=Compute.Model.ISEs(a.list, a, b, 4)
a.list <- readRDS("./simulated-data/df.list.model4_n10_c60.rds")
ise.mod4.n10.c60=Compute.Model.ISEs(a.list, a, b, 4)
a.list <- readRDS("./simulated-data/df.list.model4_n25_c30.rds")
ise.mod4.n25.c30=Compute.Model.ISEs(a.list, a, b, 4)
a.list <- readRDS("./simulated-data/df.list.model4_n25_c60.rds")
ise.mod4.n25.c60=Compute.Model.ISEs(a.list, a, b, 4)
a.list <- readRDS("./simulated-data/df.list.model4_n100_c30.rds")
ise.mod4.n100.c30=Compute.Model.ISEs(a.list, a, b, 4)
a.list <- readRDS("./simulated-data/df.list.model4_n100_c60.rds")
ise.mod4.n100.c60=Compute.Model.ISEs(a.list, a, b, 4)
save.image(file="./ISEs/Model4-ISEs.RData")
```

```{r fig.width=15,fig.height=15}
load(file="./ISEs/Model4-ISEs.RData")
out1=Get.Box.Pdf.Plots(ise.mod4.n10.c30,"n=10, c=30% (m=100)")
out2=Get.Box.Pdf.Plots(ise.mod4.n10.c60,"n=10, c=60% (m=100)")
out3=Get.Box.Pdf.Plots(ise.mod4.n25.c30,"n=25, c=30% (m=100)")
out4=Get.Box.Pdf.Plots(ise.mod4.n25.c60,"n=25, c=60% (m=100)")
out5=Get.Box.Pdf.Plots(ise.mod4.n100.c30,"n=100, c=30% (m=87)")
out6=Get.Box.Pdf.Plots(ise.mod4.n100.c60,"n=100, c=60% (m=94)")
## plot all
pdf(file="./Plots/Model4-ISEs-all.pdf", pointsize=12, paper="a4r",width=0,height=0)
grid.arrange(out1$den, out2$den, out3$den, out4$den, out5$den, out6$den, ncol=3, nrow=2)
dev.off()
```

##  Mean Integrated Squared Errors

-   Tables 2 and 3

-   MISEs

### Model 1 

```{r}
#   For n=10 only ESPA had solution
load(file="./ISEs/Model1-ISEs.RData")
c(mean(log(ise.mod1.n10.c30$espa)),var(log(ise.mod1.n10.c30$espa))^0.5/10)
c(mean(log(ise.mod1.n10.c60$espa)),var(log(ise.mod1.n10.c60$espa))^0.5/10)
c(mean(log(ise.mod1.n10.c30$km)),var(log(ise.mod1.n10.c30$km))^0.5/10)
c(mean(log(ise.mod1.n10.c60$km)),var(log(ise.mod1.n10.c60$km))^0.5/10)
c(mean(log(ise.mod1.n25.c30$espa)), mean(log(ise.mod1.n25.c30$spline)), mean(log(ise.mod1.n25.c30$km)))
c(var(log(ise.mod1.n25.c30$espa))^0.5/10, var(log(ise.mod1.n25.c30$spline))^0.5/10, var(log(ise.mod1.n25.c30$km))^0.5/10)
c(mean(log(ise.mod1.n25.c60$espa)), mean(log(ise.mod1.n25.c60$spline)), mean(log(ise.mod1.n25.c60$km)))
c(var(log(ise.mod1.n25.c60$espa))^0.5/10, var(log(ise.mod1.n25.c60$spline))^0.5/10, var(log(ise.mod1.n25.c60$km))^0.5/10)

#   Model 1: for n=100 and c=60 spline had no solution in 3 reps
c(mean(log(ise.mod1.n100.c30$espa)), var(log(ise.mod1.n100.c30$espa))^0.5/10)
c(mean(log(ise.mod1.n100.c60$km)), var(log(ise.mod1.n100.c60$km))^0.5/10)
my.mod.ise=ise.mod1.n100.c60;
x=log(my.mod.ise$spline); my.idx=!is.na(x); m=sum(my.idx);
c(mean(log(my.mod.ise$espa[my.idx])), mean(log(my.mod.ise$spline[my.idx])))
c(var(log(my.mod.ise$espa[my.idx]))^0.5/sqrt(m), var(log(my.mod.ise$spline[my.idx]))^0.5/sqrt(m))
```

### Model 2
```{r}
#   For n=10 only espa had solution.
###   (n=25,c=60): only m=35 had spline solution so compute MISE for these reps.
load(file="./ISEs/Model2-ISEs.RData")
c(mean(log(ise.mod2.n10.c30$espa)),var(log(ise.mod2.n10.c30$espa))^0.5/10)
c(mean(log(ise.mod2.n10.c60$espa)),var(log(ise.mod2.n10.c60$espa))^0.5/10)
c(mean(log(ise.mod2.n10.c30$km)),var(log(ise.mod2.n10.c30$km))^0.5/10)
c(mean(log(ise.mod2.n10.c60$km)),var(log(ise.mod2.n10.c60$km))^0.5/10)
c(mean(log(ise.mod2.n25.c30$espa)), mean(log(ise.mod2.n25.c30$spline)))
c(var(log(ise.mod2.n25.c30$espa))^0.5/10, var(log(ise.mod2.n25.c30$spline))^0.5/10)
#
my.mod.ise=ise.mod2.n25.c60;
x=log(my.mod.ise$spline); my.idx=!is.na(x); m=sum(my.idx);
c(mean(log(my.mod.ise$espa[my.idx])), mean(log(my.mod.ise$spline[my.idx])), mean(log(my.mod.ise$km[my.idx])))
c(var(log(my.mod.ise$espa[my.idx]))^0.5/sqrt(m), var(log(my.mod.ise$spline[my.idx]))^0.5/sqrt(m), var(log(my.mod.ise$km[my.idx]))^0.5/sqrt(m))
#
### Model 2: n=100 (m=83 for c=30, m=41 for c=60)
my.mod.ise=ise.mod2.n100.c30;
x=log(my.mod.ise$spline); my.idx=!is.na(x); m=sum(my.idx);
c(mean(log(my.mod.ise$espa[my.idx])), mean(log(my.mod.ise$spline[my.idx])), mean(log(my.mod.ise$km[my.idx])))
c(var(log(my.mod.ise$espa[my.idx]))^0.5/sqrt(m), var(log(my.mod.ise$spline[my.idx]))^0.5/sqrt(m), var(log(my.mod.ise$km[my.idx]))^0.5/sqrt(m))
```

### Model 3
```{r}
### Model 3: n=10 only espa had solution.
load(file="./ISEs/Model3-ISEs.RData")
c(mean(log(ise.mod3.n10.c30$espa)),var(log(ise.mod3.n10.c30$espa))^0.5/10)
c(mean(log(ise.mod3.n10.c60$espa)),var(log(ise.mod3.n10.c60$espa))^0.5/10)
c(mean(log(ise.mod3.n10.c30$km)),var(log(ise.mod3.n10.c30$km))^0.5/10)
c(mean(log(ise.mod3.n10.c60$km)),var(log(ise.mod3.n10.c60$km))^0.5/10)
c(mean(log(ise.mod3.n25.c30$espa)), mean(log(ise.mod3.n25.c30$spline)), mean(log(ise.mod3.n25.c30$km)))
c(var(log(ise.mod3.n25.c30$espa))^0.5/10, var(log(ise.mod3.n25.c30$spline))^0.5/10, var(log(ise.mod3.n25.c30$km))^0.5/10)
c(mean(log(ise.mod3.n25.c60$espa)), mean(log(ise.mod3.n25.c60$spline)), mean(log(ise.mod3.n25.c60$km)))
c(var(log(ise.mod3.n25.c60$espa))^0.5/10, var(log(ise.mod3.n25.c60$spline))^0.5/10, var(log(ise.mod3.n25.c60$km))^0.5/10)
#
### Model 3: n=100 (m=100 for c=30, m=100 for c=60)
my.mod.ise=ise.mod3.n100.c60;
x=log(my.mod.ise$spline); my.idx=!is.na(x); m=sum(my.idx);
c(mean(log(my.mod.ise$espa[my.idx])), mean(log(my.mod.ise$spline[my.idx])), mean(log(my.mod.ise$km[my.idx])))
c(var(log(my.mod.ise$espa[my.idx]))^0.5/sqrt(m), var(log(my.mod.ise$spline[my.idx]))^0.5/sqrt(m), var(log(my.mod.ise$km[my.idx]))^0.5/sqrt(m))
```

### Model 4
```{r}
### Model 4: n=10 only espa had solution.
load(file="./ISEs/Model4-ISEs.RData")
c(mean(log(ise.mod4.n10.c30$espa)),var(log(ise.mod4.n10.c30$espa))^0.5/10)
c(mean(log(ise.mod4.n10.c60$espa)),var(log(ise.mod4.n10.c60$espa))^0.5/10)
c(mean(log(ise.mod4.n10.c30$km)),var(log(ise.mod4.n10.c30$km))^0.5/10)
c(mean(log(ise.mod4.n10.c60$km)),var(log(ise.mod4.n10.c60$km))^0.5/10)
c(mean(log(ise.mod4.n25.c30$espa)), mean(log(ise.mod4.n25.c30$spline)), mean(log(ise.mod4.n25.c30$km)))
c(var(log(ise.mod4.n25.c30$espa))^0.5/10, var(log(ise.mod4.n25.c30$spline))^0.5/10, var(log(ise.mod4.n25.c30$km))^0.5/10)
c(mean(log(ise.mod4.n25.c60$espa)), mean(log(ise.mod4.n25.c60$spline)), mean(log(ise.mod4.n25.c60$km)))
c(var(log(ise.mod4.n25.c60$espa))^0.5/10, var(log(ise.mod4.n25.c60$spline))^0.5/10, var(log(ise.mod4.n25.c60$km))^0.5/10)
#
### Model 4: n=100 (m=87 for c=30, m=94 for c=60)
my.mod.ise=ise.mod4.n100.c60;
x=log(my.mod.ise$spline); my.idx=!is.na(x); m=sum(my.idx);
c(mean(log(my.mod.ise$espa[my.idx])), mean(log(my.mod.ise$spline[my.idx])), mean(log(my.mod.ise$km[my.idx])))
c(var(log(my.mod.ise$espa[my.idx]))^0.5/sqrt(m), var(log(my.mod.ise$spline[my.idx]))^0.5/sqrt(m), var(log(my.mod.ise$km[my.idx]))^0.5/sqrt(m))
```

##  Real Data Applications
```{r}
x <- citation("asaur")
x <- citation("survival")
toBibtex(x)
```

-   PLOT 1 (AML)
    -   AML, no LOGSPLINE fit...
    -   First 11 obs only (fall into 'mainitained' group)

```{r }
# pdf(file="./Plots/RealDatasets.pdf",pointsize=9,width=6,height=5)
# par(mfrow=c(2,2))
my.lwd=2
d1=aml[1:11,]
time1=d1[,1]
status1=d1[,2]
### KM
survob=Surv(time=time1,event=status1)
fit1=survfit(survob~1, conf.type="none") #conf.type="plain" for 95% CI
### ESPA
source('./Emp_SPA_Dist_Supp.R')
supp1=Emp_SPA_Dist_Supp(time1,status1)
t.grid1 = seq(supp1[1],supp1[2],by=.1)
source('./Emp_SPA_Survival_t.R')
saddle.sur1 = rep(0,length(t.grid1))
for(i in 1:length(t.grid1)){
    saddle.sur1[i] = Emp_SPA_Survival_t(t.grid1[i],time1,status1,inc=.1)
    }

plot(fit1,xlab="time", ylab="", main="Survival Function: aml (n=11)",lty=c(1,3,3), conf.int=T, yaxs="i", lwd=my.lwd, xlim=c(supp1[1],supp1[2]), col=c("black","grey","grey"))
points(t.grid1, saddle.sur1, type='l', lwd=my.lwd, col='blue', lty=2)
```

-   PLOT 2 (ovarian)
    -   n=26, no logspline solution
    
```{r}
d=ovarian
time=d$futime; status=d$fustat; #n=length(time)
survob=Surv(time=time,event=status,type='right')
fit=survfit(survob~1,conf.type="none")
source('./Emp_SPA_Dist_Supp.R')
supp=Emp_SPA_Dist_Supp(time,status)
t.grid=seq(supp[1],supp[2],by=1)
### ESPA
source('./Emp_SPA_Survival_t.R')
saddle.sur=rep(0,length(t.grid))
for(i in 1:length(t.grid)){
    saddle.sur[i] = Emp_SPA_Survival_t(t.grid[i],time,status,inc=.1)
    }

plot(fit,lty=c(1,3,3),main='ovarian (n=26)',xlab="time",
     conf.int=T, ylab="", yaxs="i", lwd=my.lwd, xlim=c(supp[1],supp[2]), col=c("black","grey","grey"))
points(t.grid,saddle.sur,lwd=my.lwd,type='l',col='blue',lty=2)
```


-   PLOT 3 (asaur)
    -   gastricXelox: n=48 and no logspline, but ESPA is good 
    
```{r}
d = gastricXelox
time=d$timeWeeks; status=d$delta; #n=length(time)
survob=Surv(time=time,event=status,type='right')
fit=survfit(survob~1,conf.type="none")

### ESPA
source('./Emp_SPA_Dist_Supp.R')
supp = Emp_SPA_Dist_Supp(time,status)
t.grid=seq(supp[1],supp[2],by=1)

source('./Emp_SPA_Survival_t.R')
saddle.sur=rep(0,length(t.grid))
for(i in 1:length(t.grid)){
    saddle.sur[i]=Emp_SPA_Survival_t(t.grid[i],time,status,inc=.1)
    }

plot(fit,lty=c(1,3,3),main='gastricXelox (n=48)',xlab="time",
     conf.int=T, ylab="",yaxs="i",lwd=my.lwd,xlim=c(supp[1],supp[2]), col=c("black","grey","grey"))
points(t.grid, saddle.sur, lwd=my.lwd, type='l', col='blue', lty=2)

```

-   PLOT 4
    -   genfan: n=70, both ESPA and logspline are similar
```{r}
d=genfan
time=d$hours; status=d$status; 
survob=Surv(time=time,event=status,type='right')
fit=survfit(survob~1,conf.type="none")
source('./Emp_SPA_Dist_Supp.R')
supp=Emp_SPA_Dist_Supp(time,status)
t.grid=seq(supp[1],supp[2],by=1)
### SPLINE
censored.obs = time[status==0]
un.censored.obs = time[status==1]
my.penalty=log(length(un.censored.obs))
fit.spl <- oldlogspline(uncensored=un.censored.obs, right=censored.obs,
                    lbound=0, penalty=my.penalty)
spline.sur=1-poldlogspline(t.grid, fit.spl)

### ESPA
source('./Emp_SPA_Survival_t.R')
saddle.sur=rep(0,length(t.grid))
for(i in 1:length(t.grid)){
    saddle.sur[i]=Emp_SPA_Survival_t(t.grid[i],time,status,inc=.1)
    }

plot(fit,lty=c(1,3,3),main='genfan (n=70)',xlab="time",
     conf.int=T, ylab="",yaxs="i",lwd=my.lwd,xlim=c(supp[1],supp[2]), col=c("black","grey","grey"))
points(t.grid,saddle.sur,lwd=my.lwd,type='l',col='blue',lty=2)
points(t.grid,spline.sur,lwd=my.lwd,type='l',col='purple',lty=4)
legend('bottomleft',legend=c('Kaplan-Meier','ESPA','SPLINE'),lwd=my.lwd,
       col=c('black','blue','purple'), lty=c(1,2,4))

#dev.off()
```

